Converging High-Throughput and High-Performance Computing: A Case Study

### 2017-04-01 overall.tex
Fix caption figure 5
Test github with Jack
Ongoing: address part of Sergey's comments, eliminated two figures

% \begin{figure}[htp]
%     \subfloat[]{
%         \includegraphics[clip,width=\columnwidth]{figures/cpu_hours.png}
%         \label{fig:core-hours-utilization}
%     }
%
%     \subfloat[]{
%         \includegraphics[clip,width=\columnwidth]{figures/backfill_consumption.png}
%         \label{fig:backfill-utilization}
%     }
% \caption{(a).(b).}
% \end{figure}

% \begin{figure}[htp]
% \includegraphics[clip,width=\columnwidth]{figures/cpu_hours.png}
% \caption{Caption.}
% \label{fig:core-hours-utilization}
% \end{figure}

@@

@@

Every detector simulation executed on Titan process 100 events. This number of
events is consistent with the physics of the use case and with the average
duration of backfill availability. The duration of a detector simulation is a
function of the number of events: the more events, the longer the simulation.
Not all events take the same time to be simulated:
Figure~\ref{fig:100event-distrib} shows a distribution of the simulation time of
1 event from $\approx$2 to $\approx$40 minutes, with a mean of $\approx$14
minutes. Considering that each worker node process up to 16 events concurrently,
100 events takes around 2 hours to process.

\begin{figure}[htp]
    \includegraphics[clip,width=\columnwidth]{figures/tx_simulation_distribution.pdf}
    \caption{Distribution of the time taken by a Geant4 detector simulation to
    simulate one event. 3000 Events; 30 Titan worker nodes; 16 simulations per
    node; 100 events per node. Average time per event $\approx$14 min. Broad
    distribution from $\approx$2 to $\approx$40 minutes.}
\label{fig:100event-distrib}
\end{figure}

@@

\begin{figure}[htp]
    \includegraphics[clip,width=\columnwidth]{figures/titan_backfill_avail.png}
    %figures/avail_time_vs_nodes_Oct_28_Nov_3_2015.pdf}
    \caption{Backfill availability measures as walltime in minutes vs number of
    work nodes during the experiment window time. Availability was measured
    62175 times: mean number of work nodes available 680; mean walltime
    available 680 minutes.}
\label{fig:backfill-distrib}
\end{figure}

@@

The AMD Opteron 6274 CPU used on Titan has 16 cores, divided into 8 compute
units. Each compute units has 1 floating point (FP) scheduler shared between 2
cores. When using 16 cores for FP-intensive calculations, each pair of cores
competes for a single FP scheduler. This creates the overhead shown in
Figure~\ref{fig:comparison-8-16cores}: the mean runtime per event for 8
concurrent simulations computing 50 events is 10.8 minutes, while for 16
simulations is 14.25 minutes (consistent with
Figure~\ref{fig:100event-distrib}). Despite an inefficiency of almost 30\%,
Titan's allocation policy based on number of worker nodes used instead of number
of cores does not justify the use of 1/2 of the cores available.

@@

The performance analysis of Titan's AMD CPUs for detector simulations helps also
to compare Titan and Grid site performances. Usually, Grid sites exposes
resources with heterogeneous CPU architectures and a maximum of 8 (virtual)
cores per worker node, while Titan's offer an homogeneous 16 cores architecture.
We used the rate of events processes per minute as a measure of the efficiency
of executing the same detector simulation on both Titan or Grid sites.
Figure~\ref{fig:comparison-titan-grid} compares the efficiencies of Titan to the
BNL and SIGNET Grid sites, normalized for 8 cores. Effective performance
per-core at Titan is $\approx$0.57 event per minute, roughly 1/2 of BNL and 1/3
of SIGNET performances.

\begin{figure}[htp]
    \includegraphics[clip,width=\columnwidth]{figures/titan_grid_event_rate.pdf}
    \caption{Comparison of the event rate for finished jobs at Titan (h1), BNL
    (h2) and GRIDNET (h3). Titan histogram shows 1/2 of the measured values to
    account for the core difference between Titan (16) and the Grid sites (8).
    Performance per core: BNL 1.126; SIGNET 0.885; Titan 0.577.}
\label{fig:comparison-titan-grid}
\end{figure}

>>>>

@@

\mtnote{Elaborate further and put it at the end, after giving several of the
measurements we would need. ``Currently, this measure of efficiency is being
refined on the base of the actual minimum amount of walltime that PanDA Broker
can request when submitting job requests. This is not a fixed value as it
depends on variable initialization and configuration overheads, simulation time,
and number of nodes obtained. Part of the undergoing effort is to enable the
measurements of  the distribution of these parameters to derive an optimization
function.''}

@@

Every detector simulation executed on Titan process 100 events. This number of
events is consistent with the physics of the use case and with the average
duration of backfill availability. The duration of a detector simulation is a
function of the number of events: the more events, the longer the simulation.
Not all events take the same time to be simulated:
% Figure~\ref{fig:100event-distrib} shows a distribution of the simulation time
% of
1 event takes from $\approx$2 to $\approx$40 minutes to be simulated, with a
mean of $\approx$14 minutes. Considering that each worker node process up to 16
events concurrently, 100 events takes around 2 hours to process.

% \begin{figure}[htp]
%     \includegraphics[clip,width=\columnwidth]{figures/tx_simulation_distribution.pdf}
%     \caption{Distribution of the time taken by a Geant4 detector simulation to
%     simulate one event. 3000 Events; 30 Titan worker nodes; 16 simulations per
%     node; 100 events per node. Average time per event $\approx$14 min. Broad
%     distribution from $\approx$2 to $\approx$40 minutes.}
% \label{fig:100event-distrib}
% \end{figure}

@@

\begin{figure}[htp]
    \includegraphics[clip,width=\columnwidth]{figures/titan_backfill_avail.png}
    %figures/avail_time_vs_nodes_Oct_28_Nov_3_2015.pdf}
    \caption{Backfill availability measures as walltime in minutes vs number of
    work nodes during the experiment window time. Availability was measured
    62175 times: mean number of work nodes available 680 (yellow line); mean
    walltime available 122 minutes (red line).}
\label{fig:backfill-distrib}
\end{figure}

@@

The AMD Opteron 6274 CPU used on Titan has 16 cores, divided into 8 compute
units. Each compute units has 1 floating point (FP) scheduler shared between 2
cores. When using 16 cores for FP-intensive calculations, each pair of cores
competes for a single FP scheduler. This creates the overhead shown in
Figure~\ref{fig:comparison-8-16cores}: the mean runtime per event for 8
concurrent simulations computing 50 events is 10.8 minutes, while for 16
simulations is 14.25 minutes (consistent with the measured distribution of the
duration of event simulation). Despite an inefficiency of almost 30\%, Titan's
allocation policy based on number of worker nodes used instead of number of
cores does not justify the use of 1/2 of the cores available.

@@

The performance analysis of Titan's AMD CPUs for detector simulations helps also
to compare Titan and Grid site performances. Usually, Grid sites exposes
resources with heterogeneous CPU architectures and a maximum of 8 (virtual)
cores per worker node, while Titan's offer an homogeneous 16 cores architecture.
We used the rate of events processes per minute as a measure of the efficiency
of executing the same detector simulation on both Titan or Grid sites.
% Figure~\ref{fig:comparison-titan-grid} compares the efficiencies of Titan to
% the BNL and SIGNET Grid sites, normalized for 8 cores. Effective performance
% per-core at Titan is $\approx$0.57 event per minute, roughly 1/2 of BNL and
% 1/3 of SIGNET performances.
Comparisons of the efficiencies of Titan to  the BNL and SIGNET Grid sites,
normalized for 8 cores, show that the effective performance per-core at Titan is
$\approx$0.57 event per minute, roughly 1/2 of BNL and  1/3 of SIGNET
performances.

% \begin{figure}[htp]
%     \includegraphics[clip,width=\columnwidth]{figures/titan_grid_event_rate.pdf}
%     \caption{Comparison of the event rate for finished jobs at Titan (h1), BNL
%     (h2) and GRIDNET (h3). Titan histogram shows 1/2 of the measured values to
%     account for the core difference between Titan (16) and the Grid sites (8).
%     Performance per core: BNL 1.126; SIGNET 0.885; Titan 0.577.\mtnote{TO BE DELETED.}}
% \label{fig:comparison-titan-grid}
% \end{figure}
